---
title: "The phenoreportr Tutorial: Development Documentation"
author: "Andrey Smelter"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The phenoreportr Tutorial: Development Documentation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Installation

First clone the repo from github:

```bash
$ git clone git@github.com:Ancestry/NatGenCOVIDpaper2022.git
$ cd clean_pheno_package/phenoreportr
```

Next, install the package from R console using `devtools` R package:

```R
> install.packages("devtools")
> library(devtools)
> devtools::install()
```

## Documentation

### Documenting R functions

We use `roxygen2` R package to document available R functions. 
In order to create documentation files for available functions first use `roxygen2` documentation style.
Invoke `devtools::document()` command for your package, this will build documentation files under `man`
directory as well as refresh `NAMESPACE` file.

```{r, engine = "bash", eval = FALSE}
> devtools::document()
Updating phenoreportr documentation
Loading phenoreportr
Writing NAMESPACE
```

### Building package documentation

We use `pkgdown` R package to build documentation for entire package.
This allow us to build static html documentation under `docs` directory
using available `*.Rd`, `*.Rmd`, and `README.md` files under 
`man` and `vignettes` directories.

First, set option to disable external CRAN lookup:

```{r, engine = "bash", eval = FALSE}
> options(pkgdown.internet = FALSE)
```

Next, invoke `pkgdown::build_site()` command to build package documentation:

```{r, engine = "bash", eval = FALSE}
> pkgdown::build_site()
Installing package into temporary library
Building pkgdown site
Reading from: '/phenoreportr'
Writing to:   '/phenoreportr/docs'
Initialising site
Building home
Building function reference
Reading 'man/extract_questions.Rd'
Reading 'man/load_data.Rd'
Reading 'man/load_ethnicity.Rd'
Reading 'man/load_pheno.Rd'
Reading 'man/load_survey.Rd'
Reading 'man/phenoreportr.Rd'
Reading 'man/phenotype_sample_size_by_response_plot.Rd'
Reading 'man/prevalence_plot.Rd'
Reading 'man/prevalence_table.Rd'
Reading 'man/response_combined_table.Rd'
Reading 'man/response_sample_size_by_question_table.Rd'
Building articles
Reading 'vignettes/development.Rmd'
Writing 'articles/development.html'
Reading 'vignettes/tutorial.Rmd'
Building news
DONE
```

### Updating package documentation

Documentation is uploaded to github pages. 
To update it commit changes in `docs` directory to master branch of `phenoreportr` package.









