---
title: "The phenoreportr Tutorial: Using phenoreportr Command Line Interface"
author: "Andrey Smelter"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The phenoreportr Tutorial: Using phenoreportr Command Line Interface}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The `phenoreportr` package provides functions to clean phenotypes and generate phenotype report(s).

## Command Line Interface

```
phenoreportr command-line interface

Generate clean phenotypes.
Generate phenotype report.

Usage:
    phenoreport.R process phenotypes --survey=<file> --pheno=<path> --trait=<name>...
    phenoreport.R process report --survey=<file> --pheno=<path> --ethnicity=<path> --trait=<name>... 
                  [--cohorts=<default>] [--format=<type>] [--dir=<path>] 
    phenoreport.R list phenotypes
    phenoreport.R -h | --help
    phenoreport.R -v | --version

Options:
    -e, --ethnicity=<path>   Path to ethnicity cohorts.
    -f, --format=<type>      The R Markdown output format to convert to [default: html_document]
    -d, --dir=<path>         The output directory for the rendered output_file.
    -p, --pheno=<path>       Phenotype path.
    -s, --survey=<file>      Path to survey file.
    -t, --trait=<name>       Name of the trait.
    -c, --cohorts=<default>  Default cohorts [default: Europe.idvs:EUR,African_American.idvs:AA,Latino.idvs:LAT,East_Asia.idvs:EAS]
    -h, --help               Show this help message.
    -v, --version            Show package version.
```

## Find `phenoreportr` CLI executable

To find `phenoreport.R` executable start your `R` shell and type:

```{r, eval = FALSE}
shQuote(system.file("exec", "phenoreport.R", package = "phenoreportr"))
```

Output:

```
"/path/to/phenoreportr/exec/phenoreport.R"
```

## Print `phenoreportr` CLI help message

```{r, engine = "bash", eval = FALSE}
Rscript /path/to/phenoreportr/exec/phenoreport.R --help
```

## Print `phenoreportr` package version

```{r, engine = "bash", eval = FALSE}
Rscript /path/to/phenoreportr/exec/phenoreport.R --version
```

## Set paths to common directories

For convenience set paths to phenotypes, ethnicity directory locations, and survey file location:

```{r, engine = "bash", eval = FALSE}
export PHENO_PATH=/path/to/phenos
export ETHNICITY_PATH=/path/to/eth
export SURVEY_FILE=/path/to/survey
export COVID_SURVEY_FILE=/path/to/covid/survey
```

## List phenotypes that are available for cleaning

To list phenotypes that are available for cleaning, run the `list phenotypes` cpmmand:

```{r, engine = "bash", eval = FALSE}
Rscript /path/to/phenoreportr/exec/phenoreport.R list phenotypes
```

This will return names of the phenotypes (and phenotype cleaning scripts) that
are currently available: 

```
[1] "COVIDHOSPITALIZED_2COVIDHOSP_1COVIDNOHOSP"                 
[2] "COVIDHOSPITALIZEDPOP_2COVIDHOSP_1POPULATION"               
[3] "COVIDSWABTEST_2COVIDPOS_1COVIDNEG"                         
[4] "COVIDSWABTESTPOP_2COVIDPOS_1POPULATION"                    
[5] "COVIDWITHPOSITIVEHOUSEMATE_2COVIDPOS_1COVIDNEG"            
[6] "COVIDWITHPOSITIVEHOUSEMATEPOP_2POPULATION_1COVIDNEGWITHEXP"
[7] "MILDASYMPTOMATICCOVID_2SYMPTOMATIC_1MILDASYMPTOMATIC"
```

## Generate clean phenotype files

To generate clean phenotype files (`.tsv` and `.pheno`), use the following command:

```{r, engine = "bash", eval = FALSE}
Rscript  /path/to/phenoreportr/exec/phenoreport.R process phenotypes \ 
-s $COVID_SURVEY_FILE -p /output/directory \
-t COVIDHOSPITALIZED_2COVIDHOSP_1COVIDNOHOSP
```

This command will generate `.tsv` and `.pheno` for `COVIDHOSPITALIZED_2COVIDHOSP_1COVIDNOHOSP`
phenotype at the specified output directory.

To generate more than one phenotype, just repeat the phenotype name:

```{r, engine = "bash", eval = FALSE}
Rscript  /path/to/phenoreportr/exec/phenoreport.R process phenotypes \ 
-s $COVID_SURVEY_FILE -p /output/directory \
-t COVIDHOSPITALIZED_2COVIDHOSP_1COVIDNOHOSP \
-t COVIDHOSPITALIZEDPOP_2COVIDHOSP_1POPULATION
```


## Generate phenotype report

To generate report for the phenotype of interest, use the following command:

```{r, engine = "bash", eval = FALSE}
Rscript path/to/phenoreportr/exec/phenoreport.R process report \
-s $SURVEY_FILE -p $PHENO_PATH -e $ETHNICITY_PATH \
-t HYPERTENSIONFAMILYHIST_2HTWITHFH_1NOHTNOFH
```

To generate reports for more than one phenotype, just repeat the phenotype name:

```{r, engine = "bash", eval = FALSE}
Rscript path/to/phenoreportr/exec/phenoreport.R process report \ 
-s $SURVEY_FILE -p $PHENO_PATH -e $ETHNICITY_PATH \
-t HYPERTENSIONFAMILYHIST_2HTWITHFH_1NOHTNOFH -t ANYNUTALLERGY_2YES_1NO
```

